### 2012-04-16 trunk/ijcis-pos-model/sources/temporaldatabases.tex
version sent to olga and nicolas.

\begin{definition}
\label{def:close-a-crisp-interval}
Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$. The CloseL$\left(I, J\right)$ function allows to close the left-open interval $I$ with respect to the first value $d$ in $J$:
\begin{align}
\mbox{CloseL} \left( I, J \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } b \neq +\infty \\
I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
\end{cases}
\end{align}
\end{definition}

Analogously, it is defined the function to close an right open interval: 
\begin{definition}
\label{def:close-a-crisp-interval-r}
CloseR$\left(I, J\right)$ 
\begin{align}
\mbox{CloseR} \left( I, J \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } a \neq -\infty \\
I=\left]d, b \right[ & \mbox{ if } a = -\infty \wedge J < I
\end{cases}
\end{align}
\end{definition}

\begin{definition}
\label{def:current-in-relation}
Consider an object $A$ and a relation $R$ in a temporal database. The fucntion Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.

@@

Now it is possible to close the current version of an object by combining the two previously defined functions.

@@

\begin{eqnarray}
%\label{eq:close-current}
\text{Close-current} \left(R, A \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ Close } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ Close } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

\subsubsection{Insert}
The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
%
%
%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used. 
There are three main cases when performing a create operation:
\begin{enumerate}
\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.

\item The object $A$ is in the relation $R$, but it is not current

\item The object $A$ is current in the relation $R$: The insert operation is rejected, a modify operation should be used.
\end{enumerate}

@@

\begin{align}
\label{eq:insert}
\INS \left(R ,A, I\right) =\\
\begin{cases}
\nonumber
R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
R \cup  \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{Delete}
The delete operation logically removes a current object $A$ which is valid in the relation $R$:
\begin{align}
\label{eq:delete}
\DEL \left(R ,A \right) =\\
\begin{cases}
\nonumber
R - \left(A, I \right), & \mbox{ if }  A \not \in R\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{Modify}
The modification of an object $A$ is usually defined by a delete operation followed by an insert operation. The modify operation is defined then by:

\begin{align}
\label{eq:modify}
\MOD \left( R, A, I \right) &=&\\
\nonumber
\INS \left( \DEL \left(R, A \right), A, I \right)
\end{align}

>>>>

\begin{definition}
\label{def:close-a-crisp-interval-r}
Consider two crisp intervals $I= \left[a ,b \right]$ and $J= \left[c ,d \right]$, $c > a$. The CloseR$\left(I, J\right)$ function allows to close the right-open interval $I$ with respect to the first value $d$ in $J$:
\begin{align}
\mbox{CloseR} \left( I, J \right) &=& \\ 
\begin{cases}
\nonumber
I & \mbox{ if } b \neq +\infty \\
I=\left[a, c \right[ & \mbox{ if } b = +\infty \wedge J > I
\end{cases}
\end{align}
\end{definition}

%Analogously, it is defined the function to close a left-open interval: 
%\begin{definition}
%\label{def:close-a-crisp-interval-l}
%CloseL$\left(I, J\right)$ 
%\begin{align}
%\mbox{CloseL} \left( I, J \right) &=& \\ 
%\begin{cases}
%\nonumber
%I & \mbox{ if } a \neq -\infty \\
%I=\left]d, b \right] & \mbox{ if } a = -\infty \wedge J < I
%\end{cases}
%\end{align}
%\end{definition}

\begin{definition}
\label{def:current-in-relation}
Consider an object $A$ and a relation $R$ in a temporal database. The function Current$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ that is current in the relation $R$, that is it, $I = \left[a, +\infty \right]$.

@@

%\begin{definition}
%\label{def:first-in-relation}
%Analogously, the function First$\left(R, A \right)$ obtains the tuple $\left(A, I \right)$ where $I$ is a right-open interval and which is the first version in the relation $R$, that is it, $I = \left[-\infty, b \right]$.
%\begin{align}
%\label{eq:first-in-relation}
%\mbox{First} \left(R, A \right) &=& \\ 
%\begin{cases}
%\nonumber
%I & \mbox{ if } \exists b \in I : (A,I) \in R \wedge a = -\infty \\
%\emptyset & \mbox{ in any other case }
%\end{cases}
%\end{align}
%
%\end{definition}

Now it is possible to close the current version of an object by using \eqref{def:close-a-crisp-interval-r} and \eqref{eq:current-in-relation}

@@

\begin{eqnarray}
%\label{eq:close-current}
\text{Close-current} \left(R, A, J \right) =\\
\begin{cases}
\nonumber
\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseR } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
\nonumber
\mbox{ if } I = \mbox{ Current } \left(R, A \right) \wedge J > \mbox{ CloseR } \left(I, J \right)   \\
\nonumber R , \text{ in any other case}
\end{cases}
\end{eqnarray}
\end{definition}

%It is also possible to append at the beggining by using \eqref{def:close-a-crisp-interval-l} and \eqref{def:first-in-relation}.
%\begin{definition}
%\label{def:append-first-version}
%Consider an object $A$, the relation $R$ and a time interval $J$. Then, the function append-first$\left(R,A,J \right)$ closes any right open version of the object $A$ if it exists and add the new version $\left(A, J \right)$.
%
%\begin{eqnarray}
%%\label{eq:close-current}
%\text{Append-first} \left(R, A, J \right) =\\
%\begin{cases}
%\nonumber
%\big \lbrace R - \left(A, I \right) \cup \left \lbrace \left(A, \mbox{ CloseL } \left(I, J\right) \right) \cup \left(A, J\right)\right \rbrace  \big \rbrace \\
%\nonumber
%\mbox{ if } I = \mbox{ First } \left(R, A \right) \wedge J < \mbox{ CloseL } \left(I, J \right)   \\
%\nonumber R , \text{ in any other case}
%\end{cases}
%\end{eqnarray}
%\end{definition}

\subsubsection{\label{subsubsec:insert}Insert}
The user wants to store an object $A$ which is valid in the relation $R$ during the time interval specified by $I = \left(a, b \right)$.
%
%
%The interpretation of the insert operation is the following: The user wants to store in the database that an object $A$  is true for some period(s) of time $I$ in the given relation $R$. To indicate that the object $A$ is current in the relation, the special value Until Current, $\UC$  is used. 
There are two main cases when performing a create operation:
\begin{enumerate}
\item The object $A$ was never in the relation $R$: The object is added with the valid-time indicated by the crisp interval $I$.

\item The object $A$ is in the relation $R$. Depending on the value of the time interval $I$, there are two possibilities:
	\begin{enumerate}
	\item Insert $\left(A, I\right)$ in the relation $R$. If the time interval $I$ does not overlap any other valid-time interval in the relation $R$.
	\item Reject the insertion, if the time interval $I$ do overlap any existing valid-time interval for the object $A$ in the relation $R$.
	\end{enumerate}

\end{enumerate}

@@

\begin{align}
\label{eq:insert}
\INS \left(R ,A, I\right) =\\
\begin{cases}
\nonumber
R \cup \left(A, I \right), & \mbox{ if }  A \not \in R\\
R \cup   \left(A, I \right) & \mbox{ if }  A \in R \wedge \\
& \big \lbrace \left( \exists J: \left(A, J \right) \in R \wedge J<I  \right) \vee \\
& \left( \exists K: \left(A, K \right) \in R \wedge K>I \vee \right) \wedge\\
&  J<I<K  \big \rbrace\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{\label{subsubsec:del}Delete}
The delete operation logically removes a current object $A$ which is valid in the relation $R$:
\begin{align}
\label{eq:delete}
\DEL \left(R ,A \right) =\\
\begin{cases}
\nonumber
R - \left(A, I \right), & \mbox{ if } \left( A, I \right) \in R\\
R, & \mbox{ otherwise }  
\end{cases} 	
\end{align}

\subsubsection{\label{subsubsec:modify}Modify}
This operation adds new information about an existing object $A$ in the relation $R$. The modify operation does not remove any previous value of the object $A$.
Note that the modify operation is only applicable when the object $A$ is current in the relation $R$; $ \left(A, I \right) \in R, I = \left[a, +\infty \right]$.

\begin{align}
\label{eq:modify}
\MOD \left( R, A, I \right) =\\
\begin{cases}
\nonumber
\mbox{ Close-current }\left(R, A, \mbox{Current}\left(R, A \right) \right) \\
 \mbox{ if Current}\left(R ,A \right) \neq \emptyset \\
 R,  \mbox{ otherwise }
\end{cases}
\end{align}

\subsubsection{\label{subsubsec:revise}Revise}
This operation changes the value of the object $A$ in the relation $R$, at a given time interval $I$. This operation is defined when a correction on the values for the object $A$ should be done.
\begin{align}
\label{eq:revise}
\mbox{ Revise}\left(R, A, I \right)=\\
\begin{cases}
\nonumber
 R - \left(A', I \right) \cup \left(A, I \right)  \\
  \mbox{ if } \left(A', I \right) \in R: A = \left(a_1, \ldots, a_n \right),\\
A'= \left(a_1', \ldots, a_n' \right)   \\   
R,  \mbox{ otherwise }
\end{cases}
\end{align}
